Hierarchical Controller Architecture in Software Defined Networks

### 2013-12-16 paper/paper.tex
minor changes

\subsection{One switch abstraction}
\label{subsec:aggr}
A controller aggregates its underlying network and behaves as one logical switch, with huge number of ports. This logical switch supports control plane protocols such as OpenFlow\cite{openflow,of-spec}, and can be controlled by an external entity (referred to as parent controller) to install forwarding rules. One switch abstraction has two parts, (i) compressing the underlying network state to represent a single switch, and (ii) translating instructions (forwarding rules) from the abstract switch to the underlying network. We present these separately below.
\subsubsection{Compression}
The primary switch abstraction comes by compressing the network into one node. Given an underlying network for a controller, the abstract switch exposed by it contains all the ports which are connected to switches outside the network. Such ports are referred to as \emph{external} ports. Conversely, ports which are connected to ports in the same network are not exposed (hidden away from abstraction) and are referred to as \emph{internal} ports. \emph{External} ports form the \emph{ingress} and \emph{egress} ports for the abstract switch. The switch abstraction maintains a list of these abstract ports, which can be (uniquely) mapped to the ports of the underlying network. This basic abstraction is sufficient for a simple control application, but we require more features for a rich and complete experience. A few such abstractions are presented below.
\begin{enumerate}
    \item \emph{Host compression}:
    Hosts in the underlying network are considered as external entities, which are exposed as hosts connected to the abstract switch.Ports which connect to hosts are considered to be\emph{external} ports. However, if the number of hosts in the underlying network is too large, exposing a list of all hosts could result in inefficient compression. Host compression can be performed in two ways, depending on the host IP distribution and the network being considered.
    \begin{itemize}
        \item \emph{IP compression} -
        Similar to the forwarding rules in an Internet router, if the IP addresses of the hosts are properly prefixed, a controller can compress host IPs and only provide prefixes to its parent controller. This works well only if the assignment of switches to controllers is uniform and in accordance with prefixes.
        \item \emph{Abstract host list} -
        A controller can expose an abstract host or list of hosts to its parent controller. The parent controller places rules for the abstract hosts in the peer controllers. This approach would be similar to LAN, and requires address translation while entering or leaving the underlying network. This is more flexible, but complicated, and can be used only when IP prefix compression fails.
    \end{itemize}
    \item \emph{Abstract flow table}:
        The basic flow table in a controller is comprise of the rules placed by the parent controller. However, for a full OpenFlow functionality, we need to support counters and queries, as in OpenFlow switches. These responses, again, must be an aggregation of the network. To this end, we must be able to efficiently represent the flows from the underlying network and obtain statistics which can be exposed to the parent. We propose a data structure to maintain, aggregate and update view of the flows in section~\ref{subsubsec:aggr-data}.
\end{enumerate}

@@

\subsubsection{Data structures}
\label{subsubsec:aggr-data}
In this section, we present data structures, maintained by a controller to translate instructions from parent controller to children switches and to translate network updates and messages from children switches to parent controller efficiently. Note that this is just one of the way in which \emph{translation} can be enforced, which we have used in our implementation. We represent the underlying network as a graph $G$ with dummy external nodes connected to each of the exposed ports. For each external node $e$, we create and maintain a directed flow graph $F_e$ rooted at $e$ as an overlay graph on $G$. $e$ here represents the \emph{ingress} port. End hosts and other external nodes act as sinks for these flows. The external nodes represent the \emph{egress} ports.
For each pair of external nodes $e_{in}$ and $e_{out}$, we can create an aggregated flow table entry based on the flow graph $F_{e_{in}}$ for $e_{in}$. We propose this method to account for possible modifications in the packet headers along the paths (multiple paths are common based on routing policies) from the \emph{ingress} port to the \emph{egress} port which could prevent us from identifying same flows by just observing flows at the \emph{ingress} and \emph{egress} ports.
Other data structures which would aid to simplify updates and aggregation are
\begin{enumerate}
    \item \emph{Path Information} - A list of $n$ best paths for some definition of best. We consider paths with least hops, but this can be varied depending on the routing policies. This can be used to choose the best path to route from between an \emph{ingress} and \emph{egress} port based on routing policies.
    \item \emph{Link Information} - A list of pointers to the paths this link belongs to, and a list of flows that pass through this link. This can be used to obtain link utilizations to select routes with minimum maximum link utilization.
\end{enumerate}
One main difference between an abstract switch and an actual switch is the inequality of costs between two exposed ports. If parameters like hop counts are being considered, the internal routing costs must also be considered. This can be achieved by adding extra \emph{attributes} to pairs of external ports which can signify the internal costs which could be considered by the parent controller. This extra feature helps reduce the sub-optimality of this solution. However, it comes at the expense of needing to update costs of each pair of nodes on network changes.

@@

\subsection{System model}
There are multiple peer controllers, managing different parts of network and keeping a global view of entire network. Network states are replicated between controllers, with some replication lag i.e. at a given instant of time network state of two controllers may be different or inconsistent. Network state could include a range of characteristics of underlying network including, topology, link utilizations, routing tables on switches. However, for the purpose of this section, we will assume that network state only consists of link utilizations. Network state on a given controller consists of latest data for links that belong to it's domain, while it may contain stale data for other links. The topology is assumed to be static and is assumed to be known by all controllers. Source routing is used to route packets/flows to destination: when a new flow arrives at the ingress switch, it is sent to the parent controller, which calculates the route using control application and network state (at that point of time). This route is marked on the packet/flow.

>>>>

\subsection{One switch abstraction}
\label{subsec:aggr}
A controller aggregates its underlying network and behaves as one logical switch, with huge number of ports. This logical switch supports control plane protocols such as OpenFlow\cite{openflow,of-spec}, and can be controlled by an external entity (referred to as parent controller) to install forwarding rules. One switch abstraction has two parts, (i) compressing the underlying network state to represent a single switch, and (ii) translating instructions (forwarding rules) from the abstract switch to the underlying network. We present these separately below.
\subsubsection{Compression}
The primary switch abstraction comes by compressing the network into one node. Given an underlying network for a controller, the abstract switch exposed by it contains all the ports which are connected to switches outside the network. Such ports are referred to as \emph{external} ports. Conversely, ports which are connected to ports in the same network are not exposed (hidden away from abstraction) and are referred to as \emph{internal} ports. \emph{External} ports form the \emph{ingress} and \emph{egress} ports for the abstract switch. The switch abstraction maintains a list of these abstract ports, which can be (uniquely) mapped to the ports of the underlying network. This basic abstraction is sufficient for a simple control application, but we require more features for a rich and complete experience. A few such abstractions are presented below.
\begin{enumerate}
    \item \emph{Host compression}:
    Hosts in the underlying network are considered as external entities, which are exposed as hosts connected to the abstract switch. Ports which connect to hosts are considered to be \emph{external} ports. However, if the number of hosts in the underlying network is too large, exposing a list of all hosts could result in inefficient compression. Host compression can be performed in two ways, depending on the host IP distribution and the network being considered.
    \begin{itemize}
        \item \emph{IP compression} -
        Similar to the forwarding rules in an Internet router, if the IP addresses of the hosts are properly prefixed, a controller can compress host IPs and only provide prefixes to its parent controller. This works well only if the assignment of switches to controllers is uniform and in accordance with prefixes.
        \item \emph{Abstract host list} -
        A controller can expose an abstract host or list of hosts to its parent controller. The parent controller places rules for the abstract hosts in the peer controllers. This approach would be similar to LAN, and requires address translation while entering or leaving the underlying network. This is more flexible, but complicated, and can be used only when IP prefix compression fails.
    \end{itemize}
    \item \emph{Abstract flow table}:
        The basic flow table in a controller comprises of the rules placed by the parent controller. However, for a full OpenFlow functionality, we need to support counters and queries, as in OpenFlow switches. These responses, again, must be an aggregation of the network. To this end, we must be able to efficiently represent the flows from the underlying network and obtain statistics which can be exposed to the parent. We propose a data structure to maintain, aggregate and update view of the flows in section~\ref{subsubsec:aggr-data}.
\end{enumerate}

@@

\subsubsection{Data structures}
\label{subsubsec:aggr-data}
In this section, we present data structures, maintained by a controller to translate instructions from parent controller to child switches and to translate network updates and messages from child switches to parent controller efficiently. Note that this is just one of the ways in which \emph{translation} can be enforced, which we have used in our implementation. We represent the underlying network as a graph $G$ with dummy external nodes connected to each of the exposed ports. For each external node $e$, we create and maintain a directed flow graph $F_e$ rooted at $e$ as an overlay graph on $G$. $e$ here represents the \emph{ingress} port. End hosts and other external nodes act as sinks for these flows. The external nodes represent the \emph{egress} ports.
For each pair of external nodes $e_{in}$ and $e_{out}$, we can create an aggregated flow table entry based on the flow graph $F_{e_{in}}$ for $e_{in}$. We propose this method to account for possible modifications in the packet headers along the paths (multiple paths are common based on routing policies) from the \emph{ingress} port to the \emph{egress} port which could prevent us from identifying same flows by just observing flows at the \emph{ingress} and \emph{egress} ports.
Other data structures which would aid to simplify updates and aggregation are
\begin{enumerate}
    \item \emph{Path Information} - A list of $n$ best paths for some definition of best. We consider paths with least hops, but this can be varied depending on the routing policies. This can be used to choose the best path to route from between an \emph{ingress} and \emph{egress} port based on routing policies.
    \item \emph{Link Information} - A list of pointers to the paths this link belongs to, and a list of flows that pass through this link. This can be used to obtain link utilizations to select routes with minimum maximum link utilization.
\end{enumerate}
One main difference between an abstract switch and an actual switch is the inequality of costs between two exposed ports. If parameters like hop counts are being considered, the internal routing costs must also be considered. This can be achieved by adding extra \emph{attributes} to pairs of external ports which can signify the internal costs which could be considered by the parent controller. This extra feature helps reduce the sub-optimality of this solution. However, it comes at the expense of needing to update costs of each pair of nodes on network changes.

@@

\subsection{System model}
There are multiple peer controllers, managing different parts of the network and keeping a global view of the entire network. Network states are replicated between controllers, with some replication lag i.e. at a given instant of time network state of two controllers may be different or inconsistent. Network state could include a range of characteristics of underlying network including, topology, link utilizations, routing tables on switches, etc. However, for the purpose of this section, we will assume that network state only consists of link utilizations. Network state on a given controller consists of latest data for links that belong to it's domain, while it may contain stale data for other links. The topology is assumed to be static and is assumed to be known by all controllers. Source routing is used to route packets/flows to destination: when a new flow arrives at the ingress switch, it is sent to the parent controller, which calculates the route using control application and network state (at that point of time). This route is marked on the packet/flow.
